### 2017-06-06 Docs/Thesis/00_Acronyms.tex
New testes scripts in Python and Bash to calc the costFunction of the prototype. Also, at the folder Docs, the thesis text was added, i'm editing it using texstudio.

% For loading this file the following package and lines are 
% required in main latex file:
%
%   \usepackage[acronym]{glossaries}
%   \loadglsentries{acronyms}
%   \makeglossaries
%
% The entries follows are according to the line:
% \newacronym{<label>}{<abbrv>}{<full description>}
%[A]
%\newglossaryentry{IDS}{name=IDS, description={Coordinated Universal Time}}
%\newglossaryentry{adt}{name=ADT, description={Atlantic Daylight Time}}
%\newglossaryentry{est}{name=EST, description={Eastern Standard Time}}
%\printglossaries
%\newglossaryentry{CLI}
%{
%    name=CLI,
%    description={Command Line Interface}
%}
%\newglossaryentry{XSS}
%{
%    name=XSS,
%    description={Type of computer security vulnerability typically found in Web applications}
%}
\newglossaryentry{botnet}
{
    name=botnet,
    description={inter-connected computers and controlled by a malicious user in order to launch repetitive tasks against certain target}
}
\newglossaryentry{container}
{
	name=container,
	description={Lightweight virtualization method for running multiple isolated instances of an single Linux operating system}
}
\newglossaryentry{testbed}
{
    name=TESTBED,
    description={Platform for testing of containers, applications and servers}
}
\newglossaryentry{intelligence}
{
    name=intelligence,
    description={Information previously processed by intelligence analysts and that are relevant, actionable and valuable for the organization}
}
\newglossaryentry{INTRIG}
{
    name=INTRIG,
    description={Information \& Networking Technologies Research \& Innovation Group}
}
\newglossaryentry{TEST}
{
    name=TEST,
    description={testing}
}
\newglossaryentry{malware}
{
	name=malware,
	description={This description will be write later}
}
\newglossaryentry{actionable}
{
	name=actionable,
	description={Data that must be specific enough to prompt some response, or change}
}
\newglossaryentry{REN-ISAC}
{
	name=REN-ISAC,
	description={Research and Education Networking Information Sharing and Analysis Center}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%[A]
\newacronym{AaaS}{AaaS}{ALTO-as-a-Service}
\newacronym{CDF}{CDF}{Cumulative Distribution Function}
\newacronym{ALTO}{ALTO}{Application-Layer Traffic Optimization}
\newacronym{API}{API}{Application Programming Interface}
\newacronym[\glsshortpluralkey=ASes]{AS}{AS}{Autonomous System}
\newacronym{ASN}{ASN}{Autonomous System Number}
%[B]
\newacronym{BGP}{BGP}{Border Gateway Protocol}
%[C]
\newacronym{CDN}{CDN}{Content Delivery Network}
\newacronym{CL}{CL}{Confidence Level}
\newacronym{CI}{CI}{Confidence Interval}
\newacronym{CC}{CC}{Correlation Coefficient}
%[D]
%[E]
\newacronym{EPS}{EPS}{Endpoint Property Service}
\newacronym{ECS}{ECS}{Endpoint Cost Service}
%[F]
\newacronym{FCC}{FCC}{Federal Communications Commission}
%[G]
\newacronym{GDB}{GDB}{Graph Database}
\newacronym{GTC}{GTC}{General Terms and Conditions}
%[H]
%[I]
\newacronym{IETF}{IETF}{Internet Engineering Task Force}
\newacronym{IXP}{IXP}{Internet eXchange Point}
\newacronym{ISP}{ISP}{Internet Service Provider}
\newacronym{IX.br}{IX.br}{Internet Steering Committee project in Brazil}
%[J]
%[K]
\newacronym{KP}{KP}{Knowledge Plane}
%[L]
\newacronym{LG}{LG}{Looking Glass}
%[M]
\newacronym{MBA}{MBA}{Measuring Broadband America}
%[N]
\newacronym{NAP}{NAP}{Network Access Point}
\newacronym{NOS}{NOS}{Network Operating System}
\newacronym{NSP}{NSP}{Network Service Provider}
\newacronym{NOC}{NOC}{Network Operation Center}
\newacronym{NSF}{NSF}{National Science Foundation}
%[O]
\newacronym{OS}{OS}{Operating System}
\newacronym{ODL}{ODL}{OpenDaylight}
%[P]
\newacronym{P2P}{P2P}{Peer-to-peer}
\newacronym{P4P}{P4P}{Provider Portal for Applications}
\newacronym{PID}{PID}{Provider-Defined Identifier}
%[Q]
\newacronym{QoE}{QoE}{Quality of Experience}
%[R]
\newacronym{RS}{RS}{Route Server}
\newacronym{RA}{RA}{Routing Arbiter}
%[S]
\newacronym{SDN}{SDN}{Software-Defined Networking}
\newacronym{SP}{SP}{Service Provider}
%[T]
%[U]
%[W]
\newacronym{WAN}{WAN}{Wide Area Network}
%[V]
\newacronym{VM}{VM}{Virtual Machine}
%[X]
%[Y]
%[Z]