### 2017-06-06 Docs/Thesis/docs/appendix/appendixA.tex

%\newpage
%\renewcommand{\appendixname}{Appendix}

%\appendix

\chapter{Revision of Probability}
\label{ap:revision-probability}

\section{Stochastic Process}

\section{Random variable}

\section{Probability Distribution Function (PDF)}

\section{Cumulative Distribution Function (CDF)}

\section{Expected value, Variance, Mean and Standard Deviation}

\section{Self-similarity}

\section{Heavy-tailed distributions}

\section{Hurst Exponent}

\section{Maximum likelihood function}

\section{Akaike information criterion}

\section{Bayesian information criterion}

\section{Estimation of parameters of stochastic distributions}

\subsection{Linear Regression and Gradient Descendant Algorithm}

\subsection{Maximum likelihood estimation}

\subsection{Others methods}