### 2017-06-06 Docs/Thesis/docs/appendix/appendixB.tex

%\renewcommand{\appendixname}{Appendix}

\chapter{Chapter 4: Additional Plots}