### 2017-06-09 Docs/Thesis/docs/appendix/appendixB.tex
new advancements on the thesis text

\chapter{Chapter ~\ref{ch:modeling-evaluation} Aditional Plots}
\label{ap:aditional-plots}

\section{CDFs distributions}

\begin{figure}[ht!]
	\centering
	\label{fig:aproximation-original-cdf-bigFlows}
	\subfloat[Chauchy]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Cauchy_aproximation_vs_Original_set}
	}
	\subfloat[Exponential(LR)]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Exponential_aproximation_(linear_regression)_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Exponential(Me)]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Exponential_aproximation_(mean)_vs_Original_set}
	}
	\subfloat[Normal]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Normal_aproximation_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Pareto(LR)]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Pareto_aproximation_(linear_regression)_vs_Original_set}
	}
	\subfloat[Pareto(MLH)]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Pareto_aproximation_(maximum_likehood)_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Weibull]{
		\includegraphics[width=78mm]{figures/apB/cdf/bigFlows_Logscale_-_Weibull_aproximation_vs_Original_set}
	}
	\caption{CDF functions for the approximations of \textit{bigFlows-pcap} inter  packet times, of many stochastic functions.}
\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:aproximation-original-cdf-wan}
	\subfloat[Chauchy]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Cauchy_aproximation_vs_Original_set}
	}
	\subfloat[Exponential(LR)]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Exponential_aproximation_(linear_regression)_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Exponential(Me)]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Exponential_aproximation_(mean)_vs_Original_set}
	}
	\subfloat[Normal]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Normal_aproximation_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Pareto(LR)]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Pareto_aproximation_(linear_regression)_vs_Original_set}
	}
	\subfloat[Pareto(MLH)]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Pareto_aproximation_(maximum_likehood)_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Weibull]{
		\includegraphics[width=78mm]{figures/apB/cdf/wan_Logscale_-_Weibull_aproximation_vs_Original_set}
	}
	\caption{CDF functions for the approximations of \textit{wan-pcap} inter  packet times, of many stochastic functions.}
\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:aproximation-original-cdf-lan}
	\subfloat[Chauchy]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Cauchy_aproximation_vs_Original_set}
	}
	\subfloat[Exponential(LR)]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Exponential_aproximation_(linear_regression)_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Exponential(Me)]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Exponential_aproximation_(mean)_vs_Original_set}
	}
	\subfloat[Normal]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Normal_aproximation_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Pareto(LR)]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Pareto_aproximation_(linear_regression)_vs_Original_set}
	}
	\subfloat[Pareto(MLH)]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Pareto_aproximation_(maximum_likehood)_vs_Original_set}
	}
	\hspace{0mm}
	\subfloat[Weibull]{
		\includegraphics[width=78mm]{figures/apB/cdf/lan_Logscale_-_Weibull_aproximation_vs_Original_set}
	}
	\caption{CDF functions for the approximations of \textit{lan-diurnal-pcap} inter  packet times, of many stochastic functions.}
\end{figure}

\section{QQplots}

%\begin{figure}[ht!]
%	\centering
%	\label{fig:aproximation-original-cdf-skype}
%	\subfloat[Chauchy]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Cauchy}
%	}
%	\subfloat[Exponential(LR)]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Exponential(LR)}
%	}
%	\hspace{0mm}
%	\subfloat[Exponential(Me)]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Exponential(Me)}
%	}
%	\subfloat[Normal]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Normal}
%	}
%	\hspace{0mm}
%	\subfloat[Pareto(LR)]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Pareto(LR)}
%	}
%	\subfloat[Pareto(MLH)]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Pareto(MLH)}
%	}
%	\hspace{0mm}
%	\subfloat[Weibull]{
%		\includegraphics[width=78mm]{figures/apB/qq/skype_QQplot_-_Weibull}
%	}
%	\caption{CDF functions for the approximations of \textit{skype-pcap} inter  packet times, of many stochastic functions.}
%\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:aproximation-original-cdf}
	\subfloat[Chauchy]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Cauchy}
	}
	\subfloat[Exponential(LR)]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Exponential(LR)}
	}
	\hspace{0mm}
	\subfloat[Exponential(Me)]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Exponential(Me)}
	}
	\subfloat[Normal]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Normal}
	}
	\hspace{0mm}
	\subfloat[Pareto(LR)]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Pareto(LR)}
	}
	\subfloat[Pareto(MLH)]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Pareto(MLH)}
	}
	\hspace{0mm}
	\subfloat[Weibull]{
		\includegraphics[width=78mm]{figures/apB/qq/bigFlows_QQplot_-_Weibull}
	}
	\caption{CDF functions for the approximations of \textit{bigFlows-pcap} inter  packet times, of many stochastic functions.}
\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:aproximation-original-cdf-lan}
	\subfloat[Chauchy]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Cauchy}
	}
	\subfloat[Exponential(LR)]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Exponential(LR)}
	}
	\hspace{0mm}
	\subfloat[Exponential(Me)]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Exponential(Me)}
	}
	\subfloat[Normal]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Normal}
	}
	\hspace{0mm}
	\subfloat[Pareto(LR)]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Pareto(LR)}
	}
	\subfloat[Pareto(MLH)]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Pareto(MLH)}
	}
	\hspace{0mm}
	\subfloat[Weibull]{
		\includegraphics[width=78mm]{figures/apB/qq/lan_QQplot_-_Weibull}
	}
	\caption{CDF functions for the approximations of \textit{lan-diurnal-pcap} inter  packet times, of many stochastic functions.}
\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:aproximation-original-cdf-wan}
	\subfloat[Chauchy]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Cauchy}
	}
	\subfloat[Exponential(LR)]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Exponential(LR)}
	}
	\hspace{0mm}
	\subfloat[Exponential(Me)]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Exponential(Me)}
	}
	\subfloat[Normal]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Normal}
	}
	\hspace{0mm}
	\subfloat[Pareto(LR)]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Pareto(LR)}
	}
	\subfloat[Pareto(MLH)]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Pareto(MLH)}
	}
	\hspace{0mm}
	\subfloat[Weibull]{
		\includegraphics[width=78mm]{figures/apB/qq/wan_QQplot_-_Weibull}
	}
	\caption{CDF functions for the approximations of \textit{wan-pcap} inter  packet times, of many stochastic functions.}
\end{figure}

\section{Validation}

\begin{figure}[ht!]
	\centering
	\label{fig:correlation-hurst-bigflows-pcap}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/bigFlows_Correlation}
		\label{correlation-skype-pcap}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/bigFlows_Hurst_Exponent}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/bigFlows_Mean}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/bigFlows_Standard_Deviation}
		\label{hurst-skype-pcap}
	}
	\caption{Statistical parameters of \textit{bigflows-pcap} and its approximations}
\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:correlation-hurst-lan-pcap}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/lan_Correlation}
		\label{correlation-skype-pcap}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/lan_Hurst_Exponent}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/lan_Mean}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/lan_Standard_Deviation}
		\label{hurst-skype-pcap}
	}
	\caption{Statistical parameters of \textit{lan-diurnal-pcap} and its approximations}
\end{figure}

\begin{figure}[ht!]
	\centering
	\label{fig:correlation-hurst-wan-pcap}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/wan_Correlation}
		\label{correlation-skype-pcap}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/wan_Hurst_Exponent}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/wan_Mean}
	}
	\hspace{0mm}
	\subfloat[]{
		\includegraphics[width=75mm]{figures/apB/val/wan_Standard_Deviation}
		\label{hurst-skype-pcap}
	}
	\caption{Statistical parameters of \textit{wan-pcap} and its approximations}
\end{figure}