### 2017-09-19 Docs/Thesis/docs/appendix/appendixE.tex
**IMPORTANTE MODIFICATIONS**: now the Prototypes/ dir is removed, and all simulations and validation tests are unther Tests/ and are in a Python project format

\chapter{Publications}

\section{IX DCA/FEEC/University of Campinas (UNICAMP) Workshop (EADCA)}
\includepdf[pages=-]{figures/apE/eadca-article2016.pdf}

\section{X DCA/FEEC/University of Campinas (UNICAMP) Workshop (EADCA)}
\includepdf[pages=-]{figures/apE/eadca-article2017.pdf}

\section{Generic Paper}
\includepdf[pages=-]{figures/apE/paper.pdf}