### 2017-09-26 Docs/EADCA-2016/abstract.tex
new papers on Docs EADCA-2016 and EADCA-2017, also a temp script called Tests/PrototypeDataProcessor/dataProcessor/calcCostFunction2.py to plot cost functions for trace, but still not automated :\

New emerging technologies have a larger unpredictability, compared to legacy equipment. They require a larger set of meaningful tests on many different scenarios. But, in the open source world is hard to find a single tool able to provide realism, speed, easy usage and flexibility at the same time.  Most of the tools are monolithic and devoted to specific purposes. This work presents a flexible and extensible framework which aims to decouple synthetic traffic modelling from its traffic generator engine. Through a new abstraction layer, it would become possible to use modern and throughput optimized tools to create realistic traffic, in an automated way. This enables a platform agnostic configuration and reproduction of complex scenarios via analytical models. Also we use \textit{pcap} files and live-capture to create "Compact Trace Descriptors".