### 2017-11-22 Docs/Thesis/docs/appendix/appendixE.tex
update on the thesis text

\chapter{Publications}
\label{ap:publications}

\begin{itemize}
	\item IX DCA/FEEC/University of Campinas (UNICAMP) Workshop (EADCA
	\item X DCA/FEEC/University of Campinas (UNICAMP) Workshop (EADCA)
	\item Using BIC and AIC for Ethernet traffic model selection. Is it worth?
	\item SIMITAR: realistic and autoconfigurable traffic generation
\end{itemize}

\includepdf[pages=-]{figures/apE/eadca-article2016.pdf}

\includepdf[pages=-]{figures/apE/eadca-article2017.pdf}

\includepdf[pages=-]{figures/apE/paperAicBic.pdf}

\includepdf[pages=-]{figures/apE/paperSimitar.pdf}

%\section{Generic Paper 3}
%\includepdf[pages=-]{figures/apE/paper3.pdf}