### 2017-11-22 Docs/Thesis/docs/appendix/todo.tex

\chapter{TODO List}

\section{Chapter 1 - Introduction - Introduction}

\section{Chapter 2 - Bibliographic Revision }

Liks úteis: 
http://www.backtrack-linux.org/forums/showthread.php?t=8740
https://sudks.wordpress.com/2016/04/27/tools-traffic-networking/
http://www.linuxlinks.com/article/2012042806090428/BenchmarkTools.html
http://uperf.org/

\begin{enumerate}

\item TODO \textbf{Geist}\cite{geist-paper} is an application level traffic generator for stress web servers. http://kkant.net/geist/ http://kkant.net/geist/doc/geist\_doc.pdf

\item TODO https://wiki.fd.io/view/Project\_Proposals/TRex https://github.com/cisco-system-traffic-generator/trex-core
\item TODO TFGEN/MCAST: http://www.pgcgi.com/hptools/
\item TODO Poisson traffic generator http://www.spin.rice.edu/Software/poisson\_gen/
\item TODO http://netsniff-ng.org/ (trafgen)
\item TODO Synthetic self-similar traffic generation http://glenkramer.com/trf\_research.shtml
\item  TODO https://github.com/steerapi/udpgen
\item TODO  http://nutsaboutnets.com/netstress/
\item TODO https://sourceforge.net/projects/ip-packet/
\item TODO PacGen http://sourceforge.net/projects/pacgen/ Packet Forger
\item TODO NTGen http://nssl.eew.technion.ac.il/files/Projects/NTGen\_v2.0/html/index.htm
\item TODO epb - Ethernet Package Bombardier  http://maz-programmersdiary.blogspot.com.br/2012/05/epb-ethernet-package-bombardier.html

\item (Programa não encontrado) FTP traffic generator: https://pdfs.semanticscholar.org/5756/cf54e46c38dd98c7ad0d5c50deb65a8b33a7.pdf
\item (Programa não encontrado) \textbf{NetSpec}: http://www.cs.columbia.edu/~hgs/internet/traffic-generator.html http://www.ittc.ku.edu/netspec/

\end{enumerate}